---
title: 'Homework: Multivariate Models'
output: html_document
---

```{r}
library(vegan)
data(dune)
data(dune.env)
?dune
```
1. Conduct an indirect ordination on the dune plant community. Specifically, visually examine a  Non-metricMulideminsional Scaling (NMDS) plot using the bray-curtis distance metric. Below is some code to help you develop a potential plot that emphasizes the role of the environmental variable “Moisture”. Describe how you interpret the graphic. What is the goal of creating such a plot? Does this analysis suggest any interesting findings with respect to the dune vegetation?
# We see that the majority of the data is to the left of the graph
```{r}
dune_mds = metaMDS(dune)
```

```{r}
plot(dune_mds, type='n')
text(dune_mds, 'sp', cex=.5)
```

```{r}
plot(dune_mds, type='n')
 text(dune_mds, 'sp', cex=.5)
 # generate vector of colors 
 color_vect = rev(terrain.colors(6))[-1]
 points(dune_mds, 'sites', pch=19, 
        col=color_vect[dune.env$Moisture])
 legend('topright', paste("Moisture =", 1:5, sep=''), 
        col=color_vect, pch=19)
```
#1st axis tries to explain species compostition is from different cites 
#most important horizontal
#second most important vertical
#shining a flashlight(moisture)
2. Carry out a direct ordination using CCA in order to test any potential hypotheses that you developed after examining the MDS plot. Specifically, carry out a test of the entire model (i.e., including all constrained axes) and also carry out tests at the scale of individual explanatory variables you included in your model if you included more than one variable. Plot your results.
```{r}
dune_ca = cca(dune)
```

```{r}
cca_dune = cca(dune ~ ., data=dune.env)
RsquareAdj(cca_dune, 100)
```
```{r}
anova(cca_dune, permutations = 999)
```
```{r}
anova(cca_dune, by='margin', permutations = 999)
```

```{r}
plot(cca_dune, type='n', scaling=1)
orditorp(cca_dune, display='sp', cex=0.5, scaling=1, col='blue')
text(cca_dune, display='bp', col='red')
```
3. Do your two analyses agree with one another or complement one another or do these two analyses seem to be suggesting different take home messages? Which analysis do you find to be more useful?
#from our first analysis we see that that moisture explains variation on the x axis(the most important axis) very well by showing colored  grouped together horizontally and not vertically. 

#The Second Analysis is easily understood by the distance of the arrows and distance from the y axis, 
#The ManagementSF arrow tells us that it is the varable that shows the most variations for the second most important axis
#We also see that moisture is the most important variable that represents the second most important axis

